---
title: "SpatialScan_report"
output:
    html_document:
        toc: true
        toc_float: true
        number_sections: true
        theme: paper
        smooth_scroll: true
---

<style type="text/css">
    #header{
        text-align: center;
    }
    body{
  font-family: Helvetica;
  font-size: 12pt;
}
/* Headers */
h1,h2,h3,h4,h5,h6{
  font-size: 14pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
```
```{r , echo=FALSE, message=F,warning=FALSE,error=F}

suppressMessages(library("optparse"))
suppressMessages(library("CellTrek"))
suppressMessages(library("dplyr"))
suppressMessages(library("Seurat"))
suppressMessages(library("viridis"))
suppressMessages(library("ConsensusClusterPlus"))
suppressMessages(library("tidyverse"))
suppressMessages(library('ComplexHeatmap'))
suppressMessages(library("SeuratData"))
suppressMessages(library("cowplot"))
suppressMessages(library("DT"))

Outputdir=snakemake@params$OutputDir
SpatialData=snakemake@params$Spatialdata
scRNAData=snakemake@params$scRNAData
STpdf=snakemake@params$STpdf
Coembedding=snakemake@params$Coembedding
SpatialDimPlot=snakemake@params$SpatialDimPlot
Celltype=snakemake@params$Celltype
Heatmap=snakemake@params$Heatmap
CelltypeFeatureplot=snakemake@params$CelltypeFeatureplot
CelltypeSpatialplot=snakemake@params$CelltypeSpatialplot
Gene=snakemake@params$Gene
GeneSpatial=snakemake@params$GeneSpatial
GeneSpatialModule=snakemake@params$GeneSpatialModule
```

```{r , echo=FALSE, message=F,warning=FALSE,error=F}
brain<-readRDS(SpatialData)
scRNA=readRDS(scRNAData)
```

## Integration with single-cell data

* **Visualize the ST data**

```{r , echo=FALSE, message=F,warning=FALSE}
st=brain
sc=scRNA
st <- RenameCells(st, new.names=make.names(Cells(st)))
sc <- RenameCells(sc, new.names=make.names(Cells(sc)))
Idents(sc) <- "celltype"
SpatialDimPlot(st)
ggsave(STpdf)
```

* **Cell charting using CellTrek**

```{r , echo=FALSE, message=F,warning=FALSE}
brain_traint <- CellTrek::traint(st_data=st, sc_data=sc, sc_assay='RNA', cell_names='celltype')
DimPlot(brain_traint, group.by = "type")
ggsave(Coembedding)
```

- After coembedding, we can chart single cells to their spatial locations. Here, we use the non-linear interpolation (intp = T, intp_lin=F) approach to augment the ST spots.

```{r , echo=FALSE, message=F,warning=FALSE}
brain_celltrek <- CellTrek::celltrek(st_sc_int=brain_traint, int_assay='traint', sc_data=sc, sc_assay = 'RNA', 
                                   reduction='pca', intp=T, intp_pnt=5000, intp_lin=F, nPCs=30, ntree=1000, 
                                   dist_thresh=0.55, top_spot=5, spot_n=5, repel_r=20, repel_iter=20, keep_model=T)$celltrek
#After cell charting, we can interactively visualize the CellTrek result using celltrek_vis
brain_celltrek$celltype <- factor(brain_celltrek$celltype, levels=sort(unique(brain_celltrek$celltype)))
Idents(brain_celltrek) <- "celltype"
SpatialDimPlot(brain_celltrek)
ggsave(SpatialDimPlot)
```

- Spatial-weighted gene co-expression analysis within the cell type of interest

```{r , echo=FALSE, message=F,warning=FALSE}
brain_celltrek_cellsub <- subset(brain_celltrek, subset=celltype== Celltype)
brain_celltrek_cellsub@assays$RNA@scale.data <- matrix(NA, 1, 1)
brain_celltrek_cellsub<- FindVariableFeatures(brain_celltrek_cellsub)
vst_df <- brain_celltrek_cellsub@assays$RNA@meta.features %>% data.frame %>% mutate(id=rownames(.))
nz_test <- apply(as.matrix(brain_celltrek_cellsub[['RNA']]@data), 1, function(x) mean(x!=0)*100)
hz_gene <- names(nz_test)[nz_test<20]
mt_gene <- grep('^mt-|^MT-', rownames(brain_celltrek_cellsub), value=T)
rp_gene <- grep('^Rpl|^Rps|^RPL|^RPS', rownames(brain_celltrek_cellsub), value=T)
vst_df <- vst_df %>% dplyr::filter(!(id %in% c(mt_gene, rp_gene, hz_gene))) %>% arrange(., -vst.variance.standardized)
feature_temp <- vst_df$id[1:2000]

brain_celltrek_scoexp_res_cc <- CellTrek::scoexp(celltrek_inp=brain_celltrek_cellsub, assay='RNA', approach='cc', sigm=140, avg_cor_min=.4, zero_cutoff=3, min_gen=40, max_gen=400)

brain_celltrek_k =c()
for(i in 1:length(names(brain_celltrek_scoexp_res_cc$gs))){
    data_frame=data.frame(gene=c(brain_celltrek_scoexp_res_cc$gs[[i]]), G=names(brain_celltrek_scoexp_res_cc$gs)[i])
    brain_celltrek_k=rbind(brain_celltrek_k,data_frame)
}
brain_celltrek_k=brain_celltrek_k %>%magrittr::set_rownames(.$gene) %>% dplyr::select(-1)
p1=pheatmap::pheatmap(brain_celltrek_scoexp_res_cc$wcor[rownames(brain_celltrek_k), rownames(brain_celltrek_k)], 
                   clustering_method='ward.D2', annotation_row=brain_celltrek_k, show_rownames=F, show_colnames=F, 
                   treeheight_row=10, treeheight_col=10, annotation_legend = T, fontsize=8,
                   color=viridis(10), main='spatial co-expression',filename=Heatmap)
pheatmap(brain_celltrek_scoexp_res_cc$wcor[rownames(brain_celltrek_k), rownames(brain_celltrek_k)], 
                   clustering_method='ward.D2', annotation_row=brain_celltrek_k, show_rownames=F, show_colnames=F, 
                   treeheight_row=10, treeheight_col=10, annotation_legend = T, fontsize=8,
                   color=viridis(10), main='spatial co-expression')
```

```{r , echo=FALSE, message=F,warning=FALSE}
brain_celltrek_cellsub<- AddModuleScore(brain_celltrek_cellsub, features=brain_celltrek_scoexp_res_cc$gs, name='CC_', nbin=10, ctrl=50, seed=42)
FeaturePlot(brain_celltrek_cellsub, grep('CC_', colnames(brain_celltrek_cellsub@meta.data), value=T))
ggsave(CelltypeFeatureplot)
```

```{r , echo=FALSE, message=F,warning=FALSE}
SpatialFeaturePlot(brain_celltrek_cellsub, grep('CC_', colnames(brain_celltrek_cellsub@meta.data), value=T))
ggsave(CelltypeSpatialplot)
```

```{r , echo=FALSE, message=F,warning=FALSE}
SpatialFeaturePlot(brain_celltrek_cellsub,features=Gene)
ggsave(GeneSpatial)
```