---
title: "SpatialScan_report"
output:
    html_document:
        toc: true
        toc_float: true
        number_sections: true
        theme: paper
        smooth_scroll: true
---

<style type="text/css">
    #header{
        text-align: center;
    }
    body{
  font-family: Helvetica;
  font-size: 12pt;
}
/* Headers */
h1,h2,h3,h4,h5,h6{
  font-size: 14pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
```
```{r , echo=FALSE, message=F,warning=FALSE,error=F}
suppressMessages(library("optparse"))
suppressMessages(library("CellTrek"))
suppressMessages(library("dplyr"))
suppressMessages(library("Seurat"))
suppressMessages(library("viridis"))
suppressMessages(library("ConsensusClusterPlus"))
suppressMessages(library("ComplexHeatmap"))
suppressMessages(library("tidyverse"))
suppressMessages(library("SeuratData"))
suppressMessages(library("cowplot"))
suppressMessages(library("DT"))
suppressMessages(library("patchwork"))

Outputdir=snakemake@params$OutputDir # nolint
SpatialData=snakemake@params$rawdata # nolint # nolint
scRNAData=snakemake@params$scRNAData # nolint
STpdf=snakemake@params$STpdf # nolint
Coembedding=snakemake@params$Coembedding # nolint
SpatialDimPlot=snakemake@params$SpatialDimPlot # nolint
Celltype=snakemake@params$Celltype # nolint
Heatmap=snakemake@params$Heatmap # nolint
CelltypeFeatureplot=snakemake@params$CelltypeFeatureplot # nolint
CelltypeSpatialplot=snakemake@params$CelltypeSpatialplot # nolint # nolint
Gene=snakemake@params$Gene # nolint
GeneSpatial=snakemake@params$GeneSpatial # nolint
GeneSpatialModule=snakemake@params$GeneSpatialModule # nolint # nolint
nCountSpatialBefore=snakemake@params$nCountSpatialBefore # nolint # nolint
nFeatureSpatialBefore=snakemake@params$nFeatureSpatialBefore # nolint
percentMTbefore=snakemake@params$percentMTbefore # nolint
nCountSpatialAfter=snakemake@params$nCountSpatialAfter # nolint
nFeatureSpatialAfter=snakemake@params$nFeatureSpatialAfter # nolint
percentMTAfter=snakemake@params$percentMTAfter # nolint
minCount=snakemake@params$minCount # nolint # nolint
minFeature=snakemake@params$minFeature# nolint
maxMT=snakemake@params$maxMT# nolint
umap=snakemake@params$umap # nolint
top10genes=snakemake@params$top10genes # nolint
```

```{r , echo=FALSE, message=F,warning=FALSE,error=F}
brain<-Seurat::Load10X_Spatial(SpatialData) # nolint
brain@images[["slice1"]]@coordinates[["tissue"]] <- as.integer(brain@images[["slice1"]]@coordinates[["tissue"]]) # nolint
brain@images[["slice1"]]@coordinates[["row"]] <- as.integer(brain@images[["slice1"]]@coordinates[["row"]]) # nolint
brain@images[["slice1"]]@coordinates[["col"]] <- as.integer(brain@images[["slice1"]]@coordinates[["col"]]) # nolint # nolint
brain@images[["slice1"]]@coordinates[["imagerow"]] <- as.integer(brain@images[["slice1"]]@coordinates[["imagerow"]]) # nolint
brain@images[["slice1"]]@coordinates[["imagecol"]] <- as.integer(brain@images[["slice1"]]@coordinates[["imagecol"]]) # nolint
scRNA=readRDS(scRNAData) # nolint
```

## Pre -processing

    These graphs show that the difference between molecular counts (molecular counts) is not only technical, but also depends on tissue anatomy

### Before QC

```{r , echo=FALSE, message=F,warning=FALSE}
mt <- PercentageFeatureSet(brain, pattern = "^mt-")
if(apply(mt,2,sum)==0){brain[["percent.mt"]] <- PercentageFeatureSet(brain, pattern = "^MT-")}else if(apply(mt,2,sum)!=0){brain[["percent.mt"]] <- PercentageFeatureSet(brain, pattern = "^mt-")} # nolint

plot1 <- VlnPlot(brain, features = "nCount_Spatial", pt.size = 0.1) + NoLegend()
plot2 <- SpatialFeaturePlot(brain, features = "nCount_Spatial") + theme(legend.position = "right") # nolint
p <- plot1|plot2
p
ggsave(nCountSpatialBefore,p, width = 10, height = 6)

p1 <- VlnPlot(brain, features = "nFeature_Spatial") + NoLegend() + theme(axis.text.x = element_blank()) # nolint
p2 <- SpatialFeaturePlot(brain, features = "nFeature_Spatial") + theme(legend.position = "right") # nolint
p <- p1|p2
p
ggsave(nFeatureSpatialBefore,p, width = 10, height = 6)

p1 <- VlnPlot(brain, features = "percent.mt") + NoLegend() + theme(axis.text.x = element_blank()) # nolint
p2 <- SpatialFeaturePlot(brain, features = "percent.mt") + theme(legend.position = "right") # nolint
p <- p1|p2
p
ggsave(percentMTbefore, p, width = 10, height = 6)
```

```{r , echo=FALSE, message=F,warning=FALSE}
minCount <- minCount
minCount <- as.numeric(minCount)
minFeature <- minFeature
minFeature <- as.numeric(minFeature)
maxMT <- maxMT
maxMT <- as.numeric(maxMT)
print(minCount)
print(minFeature)
print(maxMT)
brain <- subset(brain,nCount_Spatial>minCount&nFeature_Spatial>minFeature&percent.mt<maxMT)
```

### After QC

```{r , echo=FALSE, message=F,warning=FALSE}
plot1 <- VlnPlot(brain, features = "nCount_Spatial", pt.size = 0.1) + NoLegend()
plot2 <- SpatialFeaturePlot(brain, features = "nCount_Spatial") + theme(legend.position = "right") # nolint
p <- plot1|plot2
p
ggsave(nCountSpatialAfter,p, width = 10, height = 6)

p1 <- VlnPlot(brain, features = "nFeature_Spatial") + NoLegend() + theme(axis.text.x = element_blank()) # nolint
p2 <- SpatialFeaturePlot(brain, features = "nFeature_Spatial") + theme(legend.position = "right") # nolint
p <- p1|p2
p
ggsave(nFeatureSpatialAfter, p, width = 10, height = 6)

p1 <- VlnPlot(brain, features = "percent.mt") + NoLegend() + theme(axis.text.x = element_blank()) # nolint
p2 <- SpatialFeaturePlot(brain, features = "percent.mt") + theme(legend.position = "right") # nolint
p <- p1|p2
p
ggsave(percentMTAfter, p, width = 10, height = 6)
```

## Dimensionality reduction, clustering, and visualization

```{r , echo=FALSE, message=F,warning=FALSE}
set.seed(888)
brain <- SCTransform(brain, assay = "Spatial")
brain <- RunPCA(brain, assay = "SCT", verbose = FALSE)
brain <- FindNeighbors(brain, reduction = "pca", dims = 1:30)
brain <- FindClusters(brain, verbose = FALSE)
brain <- RunTSNE(brain, reduction = "pca", dims = 1:30)
brain <- RunUMAP(brain, reduction = "pca", dims = 1:30)
p1 <- DimPlot(brain, reduction = "umap", label = TRUE)
p2 <- SpatialDimPlot(brain, label = TRUE, label.size = 3)
p3 <- p1|p2
ggsave(umap, p3)
p <- SpatialDimPlot(brain, facet.highlight = TRUE, ncol = 5,
               cells.highlight = CellsByIdentities(brain))
p
```  

## Identification of Spatially Variable Features

```{r , echo=FALSE, message=F,warning=FALSE}
diff.wilcox = FindAllMarkers(brain)
all.markers = diff.wilcox %>% select(gene, everything()) %>% subset(p_val<0.05)
top10 = all.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
datatable(top10,rownames = FALSE, filter="top", options = list(pageLength = 5, scrollX=T)) # nolint
``` 

- Select the top 10 marker genes of each cluster to draw a heatmap

```{r , echo=FALSE, message=F,warning=FALSE, out.width=c('100%'), fig.show='hold'}
top10_genes = CaseMatch(search = as.vector(top10$gene), match = rownames(brain))  # nolint
p10 = DoHeatmap(brain, features = top10_genes, group.by = "seurat_clusters", group.bar = T, size = 4) # nolint
p10
ggsave(top10genes)
```

## Integration with single-cell data

* **Visualize the ST data**

```{r , echo=FALSE, message=F,warning=FALSE}
#brain <- subset(brain, idents = c(0))
st <- brain
sc <- scRNA
st <- RenameCells(st, new.names=make.names(Cells(st)))
sc <- RenameCells(sc, new.names=make.names(Cells(sc)))
Idents(sc) <- "celltype"
SpatialDimPlot(st)
ggsave(STpdf)
```

* **Cell charting using CellTrek**

```{r , echo=FALSE, message=F,warning=FALSE}
brain_traint <- CellTrek::traint(st_data=st, sc_data=sc, sc_assay='RNA', cell_names='celltype') # nolint
DimPlot(brain_traint, group.by = "type")
ggsave(Coembedding)
```

- After coembedding, we can chart single cells to their spatial locations. Here, we use the non-linear interpolation (intp = T, intp_lin=F) approach to augment the ST spots.

```{r , echo=FALSE, message=F,warning=FALSE}
brain_celltrek <- CellTrek::celltrek(st_sc_int=brain_traint, int_assay='traint', sc_data=sc, sc_assay = 'RNA',  # nolint
                                   reduction='pca', intp=T, intp_pnt=5000, intp_lin=F, nPCs=30, ntree=1000,  # nolint
                                   dist_thresh=0.55, top_spot=5, spot_n=5, repel_r=20, repel_iter=20, keep_model=T)$celltrek # nolint
brain_celltrek$celltype <- factor(brain_celltrek$celltype, levels=sort(unique(brain_celltrek$celltype))) # nolint
Idents(brain_celltrek) <- "celltype"
SpatialDimPlot(brain_celltrek)
ggsave(SpatialDimPlot)
```

- Spatial-weighted gene co-expression analysis within the cell type of interest

```{r , echo=FALSE, message=F,warning=FALSE}
brain_celltrek_cellsub <- subset(brain_celltrek, subset=celltype== Celltype)
brain_celltrek_cellsub@assays$RNA@scale.data <- matrix(NA, 1, 1)
brain_celltrek_cellsub<- FindVariableFeatures(brain_celltrek_cellsub)
vst_df <- brain_celltrek_cellsub@assays$RNA@meta.features %>% data.frame %>% mutate(id=rownames(.))
nz_test <- apply(as.matrix(brain_celltrek_cellsub[['RNA']]@data), 1, function(x) mean(x!=0)*100)
hz_gene <- names(nz_test)[nz_test<20]
mt_gene <- grep('^mt-|^MT-', rownames(brain_celltrek_cellsub), value=T)
rp_gene <- grep('^Rpl|^Rps|^RPL|^RPS', rownames(brain_celltrek_cellsub), value=T)
vst_df <- vst_df %>% dplyr::filter(!(id %in% c(mt_gene, rp_gene, hz_gene))) %>% arrange(., -vst.variance.standardized)
feature_temp <- vst_df$id[1:2000]

brain_celltrek_scoexp_res_cc <- CellTrek::scoexp(celltrek_inp=brain_celltrek_cellsub, assay='RNA', approach='cc', sigm=140, avg_cor_min=.5, zero_cutoff=5, min_gen=50, max_gen=500) # nolint

brain_celltrek_k = c()
for(i in 1:length(names(brain_celltrek_scoexp_res_cc$gs))){ # nolint
    data_frame=data.frame(gene=c(brain_celltrek_scoexp_res_cc$gs[[i]]), G=names(brain_celltrek_scoexp_res_cc$gs)[i])
    brain_celltrek_k=rbind(brain_celltrek_k,data_frame)
}
brain_celltrek_k=brain_celltrek_k %>%magrittr::set_rownames(.$gene) %>% dplyr::select(-1) # nolint
p1=pheatmap::pheatmap(brain_celltrek_scoexp_res_cc$wcor[rownames(brain_celltrek_k), rownames(brain_celltrek_k)],  # nolint
                   clustering_method='ward.D2', annotation_row=brain_celltrek_k, show_rownames=F, show_colnames=F,  # nolint
                   treeheight_row=10, treeheight_col=10, annotation_legend = T, fontsize=8, # nolint
                   color=viridis(10), main='spatial co-expression',filename=Heatmap) # nolint
pheatmap(brain_celltrek_scoexp_res_cc$wcor[rownames(brain_celltrek_k), rownames(brain_celltrek_k)],  # nolint
                   clustering_method='ward.D2', annotation_row=brain_celltrek_k, show_rownames=F, show_colnames=F,  # nolint
                   treeheight_row=10, treeheight_col=10, annotation_legend = T, fontsize=8, # nolint
                   color=viridis(10), main='spatial co-expression')
```

```{r , echo=FALSE, message=F,warning=FALSE}
brain_celltrek_cellsub<- AddModuleScore(brain_celltrek_cellsub, features=brain_celltrek_scoexp_res_cc$gs, name='CC_', nbin=10, ctrl=50, seed=42) # nolint
FeaturePlot(brain_celltrek_cellsub, grep('CC_', colnames(brain_celltrek_cellsub@meta.data), value=T)) # nolint
ggsave(CelltypeFeatureplot)
```

```{r , echo=FALSE, message=F,warning=FALSE}
SpatialFeaturePlot(brain_celltrek_cellsub, grep('CC_', colnames(brain_celltrek_cellsub@meta.data), value=T)) # nolint
ggsave(CelltypeSpatialplot)
```

```{r , echo=FALSE, message=F,warning=FALSE}
SpatialFeaturePlot(brain_celltrek_cellsub, features = Gene) # nolint
ggsave(GeneSpatial)
```
